---
title: "Data 605 Assignment 3"
author: "Shaya Engelman"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# ASSIGNMENT 3

## Problem Set 1

### (1)

What is the rank of the matrix A? $$ A = 
\begin{bmatrix}
1 & 2 & 3 & 4 \\
-1 & 0 & 1 & 3 \\
0 & 1 & -2 & 1 \\
5 & 4 & -2 & -3 \\
\end{bmatrix} $$

To find the rank of the matrix, we can perform Gaussian elimination to transform the matrix into it's row-echelon form.

First we'll do Row 2 + Row 1: $$\begin{bmatrix}
1 & 2 & 3 & 4 \\
0 & 2 & 4 & 7 \\
0 & 1 & -2 & 1 \\
5 & 4 & -2 & -3 \\
\end{bmatrix} $$

Next, we'll do Row 4 - 5\*Row 1: $$\begin{bmatrix}
1 & 2 & 3 & 4 \\
0 & 2 & 4 & 7 \\
0 & 1 & -2 & 1 \\
0 & -6 & -17 & -23 \\
\end{bmatrix} $$

Now, we'll do Row 3 - 1/2\*Row 2: $$\begin{bmatrix}
1 & 2 & 3 & 4 \\
0 & 2 & 4 & 7 \\
0 & 0 & -4 & -2.5 \\
0 & -6 & -17 & -23 \\
\end{bmatrix} $$

Row 4 + 3\*Row 2: $$\begin{bmatrix}
1 & 2 & 3 & 4 \\
0 & 2 & 4 & 7 \\
0 & 0 & -4 & -2.5 \\
0 & 0 & -5 & 5 \\
\end{bmatrix} $$

Finally, Row 4 + (-1.25\*Row 3): $$\begin{bmatrix}
1 & 2 & 3 & 4 \\
0 & 2 & 4 & 7 \\
0 & 0 & -4 & -2.5 \\
0 & 0 & -0 & 8.125 \\
\end{bmatrix} $$

Lastly, for interpretability, Row 3/-4: $$\begin{bmatrix}
1 & 2 & 3 & 4 \\
0 & 2 & 4 & 7 \\
0 & 0 & 1 & 0.625 \\
0 & 0 & -0 & 8.125 \\
\end{bmatrix} $$

All the rows are non- zero rows so the rank of the matrix is **4.**

Check our work using R:

Using R returns the same answer.

```{r}
A <- matrix(c(1, 2, 3, 4,
              -1, 0, 1, 3,
              0, 1, -2, 1,
              5, 4, -2, -3), nrow = 4, byrow = TRUE)

rank_A <- qr(A)$rank

print(rank_A)
```

### (2)

Given an $mxn$ matrix where $m > n$, what can be the maximum rank? The minimum rank, assuming that the matrix is non-zero?

The **maximum rank** for this matrix would be $n$. This is due to the maximum rank being the smaller of the rows or columns. The **minimum rank** would be **1** due to a **non-zero** matrix having at least one non-zero row or column.

### (3)

What is the rank of the matrix B? $$ B = 
\begin{bmatrix}
1 & 2 & 1 \\
3 & 6 & 3 \\
2 & 4 & 2 \\
\end{bmatrix} $$

Solving using reduced row echelon form:

Row 2 - 3\*Row 1:

$$\begin{bmatrix}
1 & 2 & 1 \\
0 & 0 & 0 \\
2 & 4 & 2 \\
\end{bmatrix}$$

Row 3 - 2\*Row 1:

$$\begin{bmatrix}
1 & 2 & 1 \\
0 & 0 & 0 \\
0 & 0 & 0 \\
\end{bmatrix}$$

Since there is only one non-zero row, the rank of the matrix is **1**

Check our work using R:

Using R returns the same answer.

```{r}
B <- matrix(c(1, 2, 1,
              3 , 6, 3,
              2, 4, 2), nrow = 3, byrow = TRUE)

rank_B <- qr(B)$rank

print(rank_B)
```

## Problem set 2

Compute the eigenvalues and eigenvectors of the matrix A. You'll need to show your work. You'll need to write out the characteristic polynomial and show your solution. $$AB = 
\begin{bmatrix}
1 & 2 & 3 \\
0 & 4 & 5 \\
0& 0 & 6 \\
\end{bmatrix} $$

To find the eigenvalues of matrix $B$, we need to solve the characteristic equation $\det(B - \lambda I) = 0$, where $\lambda$ is the eigenvalue and $I$ is the identity matrix. The characteristic equation is given by:

$$
\det(B - \lambda I) = \det
\begin{bmatrix}
    1-\lambda & 2 & 3 \\
    0 & 4-\lambda & 5 \\
    0 & 0 & 6-\lambda \\
\end{bmatrix} = 0
$$

Expanding the determinant, we get:

$$
(1 - \lambda)(4 - \lambda)(6 - \lambda) = 0
$$

So by setting the roots to zero, we can easily see the roots of the polynomial, and thus the eigenvalues of the matrix, are **1, 4 and 6.**

By plugging the eigenvalues back into the equation one at a time, we can find the eigenvectors as well.

##### For $\lambda = 1$:

$$
\begin{bmatrix}
    1-1 & 2 & 3 \\
    0 & 4-1 & 5 \\
    0 & 0 & 6-1 \\
\end{bmatrix}   = \begin{bmatrix}
    0 & 2 & 3 \\
    0 & 3 & 5 \\
    0 & 0 & 5 \\
\end{bmatrix}
$$ Divide row 1 by 2: $$
\begin{bmatrix}
    0 & 1 & 3/2 \\
    0 & 3 & 5 \\
    0 & 0 & 5 \\
\end{bmatrix}
$$ Row 2 - 3\*row 1: $$
\begin{bmatrix}
    0 & 1 & 3/2 \\
    0 & 0 & 1/2 \\
    0 & 0 & 5 \\
\end{bmatrix}
$$ Multiply row 2 by 2: $$
\begin{bmatrix}
    0 & 1 & 3/2 \\
    0 & 0 & 1 \\
    0 & 0 & 5 \\
\end{bmatrix}
$$ Row 1 - row 2(3/2): $$
\begin{bmatrix}
    0 & 1 & 0 \\
    0 & 0 & 1 \\
    0 & 0 & 5 \\
\end{bmatrix}
$$ Row 3 - row 2\*5: $$
\begin{bmatrix}
    0 & 1 & 0 \\
    0 & 0 & 1 \\
    0 & 0 & 0 \\
\end{bmatrix}
$$ $$
\begin{aligned}
x_1 &= t, \\
x_2 &= 0, \\
x_3 &= 0, \\
\mathbf{v}_1 &= \begin{bmatrix} 1 \\ 0 \\ 0 \end{bmatrix}.
\end{aligned}
$$

##### For $\lambda = 4$:

$$
\begin{bmatrix}
    1-4 & 2 & 3 \\
    0 & 4-4 & 5 \\
    0 & 0 & 6-4 \\
\end{bmatrix}   = \begin{bmatrix}
    -3 & 2 & 3 \\
    0 & 0 & 5 \\
    0 & 0 & 2 \\
\end{bmatrix}
$$ Divide row 1 by -3: $$
\begin{bmatrix}
    1 & -2/3 & -1 \\
    0 & 0 & 5 \\
    0 & 0 & 2 \\
\end{bmatrix}
$$ Divide row 2 by 5: $$
\begin{bmatrix}
    1 & -2/3 & -1 \\
    0 & 0 & 1 \\
    0 & 0 & 2 \\
\end{bmatrix}
$$ Row 1 + row 2: $$
\begin{bmatrix}
    1 & -2/3 & 0 \\
    0 & 0 & 1 \\
    0 & 0 & 2 \\
\end{bmatrix}
$$ Row 3 - row 2\*2: $$
\begin{bmatrix}
    1 & -2/3 & 0 \\
    0 & 0 & 1 \\
    0 & 0 & 0 \\
\end{bmatrix}
$$ $$
\begin{aligned}
x_1 &= 2_t/3, \\
x_2 &= t, \\
x_3 &= 0, \\
\mathbf{v}_1 &= \begin{bmatrix} 2/3 \\ 1 \\ 0 \end{bmatrix}.
\end{aligned}
$$

##### For $\lambda = 6$:

$$
\begin{bmatrix}
    1-6 & 2 & 3 \\
    0 & 4-6 & 5 \\
    0 & 0 & 6-6 \\
\end{bmatrix}   = \begin{bmatrix}
    -5 & 2 & 3 \\
    0 & -2 & 5 \\
    0 & 0 & 0 \\
\end{bmatrix}
$$ Divide row 1 by -5: $$\begin{bmatrix}
    1 & -2/5 & -3/5 \\
    0 & -2 & 5 \\
    0 & 0 & 0 \\
\end{bmatrix}
$$ Divide row 2 by -2: $$\begin{bmatrix}
    1 & -2/5 & -3/5 \\
    0 & 1 & -5/2 \\
    0 & 0 & 0 \\
\end{bmatrix}
$$ Row 1 + row 2\*(2/5): $$\begin{bmatrix}
    1 & 0 & -8/5 \\
    0 & 1 & -5/2 \\
    0 & 0 & 0 \\
\end{bmatrix}
$$ $$
\begin{aligned}
x_1 &= 8_t/5, \\
x_2 &= 5_t/2, \\
x_3 &= t, \\
\mathbf{v}_1 &= \begin{bmatrix} 8/5 \\ 5/2 \\ 1 \end{bmatrix}.
\end{aligned}
$$

Check our work using R:

Using R returns the following results. While they might seem to be different results as the above calculations, they are just normalized and are the same results.

```{r}
a = matrix(c(1,2,3,0,4,5,0,0,6), ncol = 3, nrow = 3, byrow = TRUE)

eigen(a)
```