---
title: "Untitled"
author: "Shaya Engelman"
date: "2024-03-06"
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

1.  Let X1, X2, . . . , Xn be n mutually independent random variables,
    each of which is uniformly distributed on the integers from 1 to k.
    Let Y denote the minimum of the Xi's. Find the distribution of Y .

Y is the minimum of X. Let x be the smallest number. The probability
that Y is greater than or equal to x is multiplying the probability of
each X event together. P(X1) \* P(X2) \* . . . P(Xn). Each probability
is written as $(k - x + 1)/k$. When mutiplying them all together we get
$(k - x + 1)^n/k^n$. Since we were told that $Xi$ is uniformly
distributed we are looking at 1 - the probability that Y is greater than
x, $$1 - (k - x + 1)^n/k^n$$ This is the CDF of Y. The PDF is the
derivative of the CDF. $$f(x)
= n(k - x)^{n-1}/k^n$$

We can visualize the distribution with the below plot:

```{r}
set.seed(1125)
n <- 1000
k <- 20
Y <- c()
for (i in 1:n){
  Xn <- sample(1:k, 5)
  Y <- c(Y, min(Xn))
}
ggplot(data.frame(table(Y)), aes(Y, Freq)) + 
  geom_bar(stat = "identity", fill = "skyblue2", width = 1) +
  ylab("Count") + 
  ggtitle("Distribution of Y") +
  theme(plot.title = element_text(hjust = 0.5)) 
```

2.  Your organization owns a copier (future lawyers, etc.) or MRI
    (future doctors). This machine has a manufacturer's expected
    lifetime of 10 years. This means that we expect one failure every
    ten years. (Include the probability statements and R Code for each
    part.).

**I found the wording of the problems a bit ambiguous. I solved them
assuming the problems were asking for the probability of the machine not
breaking within the first 8 years (inclusive). Not that it had broken by
then and not that they always broke in the ninth year. Just that it had
not broken within the first 8**

a.  What is the probability that the machine will fail after 8 years?.
    Provide also the expected value and standard deviation. Model as a
    geometric. (Hint: the probability is equivalent to not failing
    during the first 8 years..)

$$P(X=x) = (1-p)^{x-1} \cdot p$$

```{r}

p <- 1/10
k <- 8

geom_prob <- 1 - (1 - p)^(k - 1) * p

expected_value <- 1 / p
standard_deviation <- sqrt((1 - p) / p^2)

cat("Geometric Probability:", geom_prob, "\n")
cat("Expected value: ", expected_value, "\n")
cat("Standard Deviation: ", standard_deviation, "\n")
```

b.  What is the probability that the machine will fail after 8 years?.
    Provide also the expected value and standard deviation. Model as an
    exponential.

$$f(x; \lambda) = \lambda \cdot e^{-\lambda x}$$

```{r}
lambda <- 0.1

exp_prob <- exp(-lambda * 8)

expected_value <- 1 / lambda
standard_deviation <- 1 / lambda

cat("Exponential Probability:", exp_prob, "\n")
cat("Expected value: ", expected_value, "\n")
cat("Standard Deviation: ", standard_deviation, "\n")

```

c.  What is the probability that the machine will fail after 8 years?.
    Provide also the expected value and standard deviation. Model as a
    binomial. (Hint: 0 success in 8 years)

$$P(X = k) = \binom{n}{k} \cdot p^k \cdot (1-p)^{n-k}$$

```{r}
n <- 8
p <- 0.1
k <- 0

binom_prob <- dbinom(k, n, p)

expected_value <- n * p
standard_deviation <- sqrt(n * p * (1 - p))

cat("Binomial Probability:", binom_prob, "\n")
cat("Expected value: ", expected_value, "\n")
cat("Standard Deviation: ", standard_deviation, "\n")

```

d.  What is the probability that the machine will fail after 8 years?.
    Provide also the expected value and standard deviation. Model as a
    Poisson.

$$P(X=k) = \frac{e^{-\lambda} \cdot \lambda^k}{k!}$$

```{r}
lambda <- 0.8
k <- 0
pois_prob <- dpois(k, lambda)

lambda <- 1
expected_value <- lambda
standard_deviation <- sqrt(lambda)
cat("Poisson Probability:", pois_prob, "\n")
cat("Expected value: ", expected_value, "\n")
cat("Standard Deviation: ", standard_deviation, "\n")
```
