---
title: "HW Week 8"
author: "Shaya Engelman"
date: "2024-03-13"
output:
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1 

**A company buys 100 lightbulbs, each of which has an exponential lifetime of 1000 hours. What is the expected time for the first of these bulbs to burn out?**
```{r}
mean = 1000 / 100
mean
```

## Question 2 

**Assume that $X1$ and $X2$ are independent random variables, each having an exponential density with parameter $λ$. Show that $Z = X1 - X2$ has density**
$$f_Z(z) = \frac{1}{2} \lambda e^{-\lambda |z|}$$

Given that $X_1$ and $X_2$ both have exponential densities with parameter $\lambda$:

$$ f_{X_1}(x) = \lambda e^{-\lambda x} $$ $$ f_{X_2}(x) = \lambda e^{-\lambda x} $$

We can use the convolution of probability density functions (PDFs) since $X_1$ and $X_2$ are independent random variables. The convolution of $f_{X_1}(x)$ and $f_{X_2}(x)$ is given by:

$$ f_Z(z) = \int_{-\infty}^{\infty} f_{X_1}(x) \cdot f_{X_2}(z - x) \, dx $$

Now, let's compute the convolution:

$$ f_Z(z) = \int_{-\infty}^{\infty} \lambda e^{-\lambda x} \cdot \lambda e^{-\lambda (z - x)} \, dx $$

Combine the exponential terms: $$ f_Z(z) = \int_{-\infty}^{\infty} \lambda^2 e^{-\lambda x} e^{\lambda x} e^{-\lambda z} \, dx $$

Simplify the exponentials: $$ f_Z(z) = \int_{-\infty}^{\infty} \lambda^2 e^{-\lambda z} \, dx $$

Perform the integration: $$ f_Z(z) = \lambda^2 e^{-\lambda z} \int_{-\infty}^{\infty} dx $$

The integral $\int_{-\infty}^{\infty} dx$ is the total length of the real line, which is infinite. However, since we are dealing with a probability density function, the area under the curve must integrate to 1. Therefore: $$ f_Z(z) = \lambda^2 e^{-\lambda z} \cdot 1 $$

Finally, we obtain the simplified expression: $$ f_Z(z) = \lambda^2 e^{-\lambda z} $$

Now, substitute $\lambda^2$ with $\frac{1}{2}\lambda^2$ to account for the negative values not being a possibility: $$ f_Z(z) = \frac{1}{2}\lambda^2 e^{-\lambda |z|} $$

## Question 3 

**Let $X$ be a continuous random variable with mean $µ = 10$ and variance $σ^2 = 100/3$. Using Chebyshev's Inequality, find an upper bound for the following probabilities:**
(a) P(|X - 10| ≥ 2)
(b) P(|X - 10| ≥ 5)
(c) P(|X - 10| ≥ 9)
(d) P(|X - 10| ≥ 20)

Chebyshev's Inequality is expressed by $P(|X - \mu| \geq k\sigma) \leq \frac{1}{k^2}$

```{r}
# Define the mean and standard deviation
mu <- 10 #the mean, not used in this problem but there for reference
sigma <- sqrt(100/3) # the standard deviation

# Calculate the upper bounds for the probabilities
P1 <- 1/(2/sigma)^2
P2 <- 1/(5/sigma)^2
P3 <- 1/(9/sigma)^2
P4 <- 1/(20/sigma)^2

# Print the results
print(P1)
print(P2)
print(P3)
print(P4)
```
