---
title: "Untitled"
author: "Shaya Engelman"
date: \today
output:
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1

**Use integration by substitution to solve the integral below:**
$$∫4e-^{7x}dx$$

$$u = -7x$$ $$du = -7dx$$ $$dx = \frac{du}{-7}$$

$$∫4e^u\frac{-1}{7} du$$ $$\frac{4}{-7} ∫e^u du$$
$$\frac{4}{-7} e^u + C$$ $$\frac{4}{-7} e^{-7x} + C$$

## 2

**Biologists are treating a pond contaminated with bacteria. The level
of contamination is changing at a rate of dNdt=−3150t4−220 bacteria per
cubic centimeter per day, where t is the number of days since treatment
began. Find a function N( t ) to estimate the level of contamination if
the level after 1 day was 6530 bacteria per cubic centimeter**

$$∫(-3150t^4 - 220) dt$$ $$-630t^5 - 220t + C$$

$$N( t ) = -630t^5 - 220t + C$$ $$6530 = -630(1)^5 - 220(1) + C$$
$$6530 = -630 - 220 + C$$ $$6530 = -850 + C$$ $$C = 7380$$

$$N( t ) = -630t^5 - 220t + 7380$$

## 3

**Find the total area of the red rectangles in the figure below, where
the equation of the line is** $f(x) = 2x - 9$.

There are 4 rectangles in the plot of this function within the closed
interval [4.5,8.5]. The width of each rectangle is 1. The height of each
rectangle is the value of the function at the midpoint of the rectangle.

That leads us to get the following: $$\int_{4.5}^{8.5} (2x - 9) \, dx$$
$$[x^2 - 9x]_{4.5}^{8.5}$$ $$(8.5^2 - 9(8.5)) - (4.5^2 - 9(4.5))$$
$$(72.25 - 76.5) - (20.25 - 40.5)$$ $$-4.25 - (-20.25)$$ $$16$$ **16**

## 4

**Find the area of the region bounded by the graphs of the given
equations.** $y = x^2 - 2x - 2$, $y = x + 2$

In order to find the area of the region bounded by the graphs of the
given equations, we need to find the points of intersection of the two
functions. We can do this by setting the two equations equal to each
other and solving for x. $x^2 - 2x - 2 = x + 2$ $x^2 - 3x - 4 = 0$
$(x - 4)(x + 1) = 0$ $x = 4$ or $x = -1$

The area of the region bounded by the graphs of the given equations is
given by the integral of the difference of the two functions over the
interval from -1 to 4. $$\int_{-1}^{4} (x+2)-(x^2-2x-2) \, dx$$
$$ = \int_{-1}^{4} (x+2-x^2+2x+2) \, dx$$
$$ = \int_{-1}^{4} (-x^2+3x+4) \, dx$$

We'll use R to calculate the integral.

```{r}
# Define the integrand
integrand <- function(x) {
  -x^2 + 3*x + 4
}

# Compute the integral from -1 to 4
result <- integrate(integrand, -1, 4)

# Print the result
print(result$value)
```

The area of the region bounded by the graphs of the given equations is
**20.833**

## 5

**A beauty supply store expects to sell 110 flat irons during the next
year. It costs \$3.75 to store one flat iron for one year. There is a
fixed cost of \$8.25 for each order. Find the lot size and the number of
orders per year that will minimize inventory costs.**

The ordering cost ($DS/Q$) is the cost associated with placing orders.
Each order has a fixed cost $S$, and if you need $D$ units per year,
then you will need to place $D/Q$ orders per year. So, the total
ordering cost per year is $DS/Q$.

The holding cost ($QH/2$) is the cost to hold one unit in inventory for
one year. If you order $Q$ units each time, then on average (assuming
constant demand and no shortages), you will have $Q/2$ units in
inventory. So, the total holding cost per year is $QH/2$.

The total cost $C$ is the sum of the ordering cost and the holding cost:

$$
C = \frac{DS}{Q} + \frac{QH}{2}
$$

To find the minimum of this function, we can take the derivative of $C$
with respect to $Q$, set it equal to zero, and solve for $Q$..

Take the derivative of $C$ with respect to $Q$: The derivative of $C$
with respect to $Q$ is $-DS/Q^2 + H/2$.

$$
\frac{dC}{dQ} = -\frac{DS}{Q^2} + \frac{H}{2}
$$

Set the derivative equal to zero and solve for $Q$: Setting the
derivative equal to zero gives us the equation $-DS/Q^2 + H/2 = 0$.

$$
-\frac{DS}{Q^2} + \frac{H}{2} = 0
$$

Multiplying through by $Q^2$ gives:

$$
-DS + \frac{HQ^2}{2} = 0
$$

Rearranging terms gives:

$$
\frac{HQ^2}{2} = DS
$$

Multiplying through by 2 gives:

$$
HQ^2 = 2DS
$$

Finally, solving for $Q$ gives:

$$
Q = \sqrt{\frac{2DS}{H}}
$$

This is the calculus way of finding the order quantity that minimizes
the total cost. This quantity is known as the Economic Order Quantitiy,
often abbreviated as EOQ, and is the standard formula used in inventory
management.

Let's calculate the EOQ and the number of orders per year using the
given values of $D$, $S$, and $H$.

$$EOQ = \sqrt{\frac{2 \times 110 \times 8.25}{3.75}} $$
$$= \sqrt{\frac{1815}{3.75}} $$
$$= \sqrt{484} $$
$$= 22$$

To find the optimal amount of orders, we divide the annual demand by the
EOQ:

$$\text{Number of Orders per Year} = \frac{D}{EOQ}$$
$$= \frac{110}{22} $$
$$= 5$$

So the optimal order quantity (EOQ) is **22** flat irons, and the number
of orders per year is **5**.

We can check our work using R:

```{r}
# Define the parameters
D <- 110      # Annual demand
S <- 8.25     # Cost per order
H <- 3.75     # Holding cost per unit per year

# Calculate EOQ
EOQ <- sqrt((2 * D * S) / H)

# Calculate the number of orders per year
num_orders <- D / EOQ

# Print the results
print(paste("Optimal order quantity (EOQ):", EOQ))
print(paste("Number of orders per year:", num_orders))
```

## 6

**Use integration by parts to solve the integral below.**
$$∫ln( 9x ) · x^6 dx$$

The formula for integration by parts is:

$$∫u dv = uv - ∫v du$$

In this case, we can choose $u = ln(9x)$ and $dv = x^6 dx$. Then, we need to find $du$ and $v$.

$$u = ln(9x)$$
$$du = \frac{1}{9x} · 9 dx$$
$$du = \frac{1}{x} dx$$

$$dv = x^6 dx$$
$$v = \frac{1}{7} x^7$$

Now, we can apply the formula for integration by parts:

$$∫ln(9x) · x^6 dx = ln(9x) · \frac{1}{7} x^7 - ∫\frac{1}{x} · \frac{1}{7} x^7 dx$$
$$= \frac{1}{7} x^7 ln(9x) - \frac{1}{7} ∫x^6 dx$$
$$= \frac{1}{7} x^7 ln(9x) - \frac{1}{7} \frac{1}{7} x^7 + C$$
$$= \frac{1}{7} x^7 ln(9x) - \frac{1}{49} x^7 + C$$

So, the solution to the integral is:
$$\frac{1}{7} x^7 ln(9x) - \frac{1}{49} x^7 + C$$

## 7

**Determine whether f ( x ) is a probability density function on the interval [1, e^6] . If not, determine the value of the 
definite integral.**
$$f(x) = \frac{1}{6x}$$

To determine if $f(x)$ is a probability density function on the interval $[1, e^6]$, we need to check if the integral of $f(x)$ over the interval equals 1.

$$∫_{1}^{e^6} \frac{1}{6x} dx$$

$$= \frac{1}{6} ∫_{1}^{e^6} \frac{1}{x} dx$$

$$= \frac{1}{6} [ln(x)]_{1}^{e^6}$$

$$= \frac{1}{6} [ln(e^6) - ln(1)]$$

$$= \frac{1}{6} [6 - 0]$$

$$= 1$$

Since the integral of $f(x)$ over the interval $[1, e^6]$ equals 1, $f(x)$ is a probability density function on that interval.
