---
title: "HW 15"
author: "Shaya Engelman"
date: \today
output:
  pdf_document:
    latex_engine: xelatex
---

# 1.

Find the equation of the regression line for the given points.
Round any final values to the nearest hundredth, if necessary.
$$ ( 5.6, 8.8 ), ( 6.3, 12.4 ), ( 7, 14.8 ), ( 7.7, 18.2 ), ( 8.4, 20.8 ) $$

```{r}
x <- c(5.6, 6.3, 7, 7.7, 8.4)
y <- c(8.8, 12.4, 14.8, 18.2, 20.8)

lm(y ~ x)
```

The equation of the regression line is $y = 4.26x - 14.8$


# 2.

Find all local maxima, local minima, and saddle points for the function given below.
Write your answer(s) in the form $( x, y, z )$.
Separate multiple points with a comma.
$$ f ( x, y ) = 24x - 6xy^2 - 8y^3 $$

To solve this we find the critical points by taking the partial derivatives of the function and setting them equal to zero.
$$ f_x = 24 - 6y^2 = 0 $$
$$ 24 = 6y^2 $$
$$ y^2 = 4 $$
$$ y = \pm 2 $$

$$ f_y = -12xy - 24y^2 = 0 $$
$$ -12xy - 24y^2 = 0 $$
$$ -12x - 24y = 0 $$
$$ -12x = 24y $$
$$ x = -2y $$
$$ x = -2(\pm 2) $$
$$ x = \pm 4 $$

Now we plug the critical points into the original function to find the $z$ values.

```{r}
equation = function(x,y){
  z = 24*x - 6*x*y^2 - 8*y^3
  return(c(x,y,z))
}
print(rbind(equation(-4,2),equation(4,-2)))
```

So the critical points are $(-4, 2, -64)$ and $(4, -2, 64)$.

Now we need to find the second partial derivatives to determine the nature of the critical points.

$ f_{xx} = 0$ and $f_{yy} = -12x - 48y$ and $f_{xy} = -12y$
$$ D = f_{xx}f_{yy} - f_{xy}^2 $$
$$ D = 0(-12x - 48y) - (-12y)^2 $$
$$ D = 144y^2 $$
Since $D < 0$, the critical points are saddle points.

# 3.

A grocery store sells two brands of a product, the "house" brand and a "name" brand.
The manager estimates that if she sells the "house" brand for $x$ dollars and the "name" brand for $y$ dollars, she will be able to sell $81 - 21x + 17y$ units of the "house" brand and $40 + 11x - 23y$ units of the "name" brand.

**Step 1.** Find the revenue function $R ( x, y )$.

The revenue function is given by the product of the price and the quantity sold.
$$ R ( x, y ) = x(81 - 21x + 17y) + y(40 + 11x - 23y) $$
$$ R ( x, y ) = 81x - 21x^2 + 17xy + 40y + 11xy - 23y^2 $$
$$ R ( x, y ) = -21x^2 + 28xy - 23y^2 + 81x + 40y $$


**Step 2.** What is the revenue if she sells the "house" brand for $2.30 and the "name" brand for $4.10?
```{r}
x <- 2.3
y <- 4.1

revenue <- - (21 * x^2) + (28 * y * x) - (23 * y^2) + (81 * x) + (40 * y) 
cat("The revenue is $", revenue, "\n")
```

# 4.

A company has a plant in Los Angeles and a plant in Denver.
The firm is committed to produce a total of 96 units of a product each week.
The total weekly cost is given by $C(x, y) = \frac{1}{6} x^2 + \frac{1}{6} y^2 + 7x + 25y + 700$, where $x$ is the number of units produced in Los Angeles and $y$ is the number of units produced in Denver.
How many units should be produced in each plant to minimize the total weekly cost?

This is a univariate function.
To find the minimum, we take the partial derivatives of the function and set them equal to zero.
$$ C_x = \frac{1}{3} x + 7 = 0 $$
$$ \frac{1}{3} x = -7 $$
$$ x = -21 $$
$$ C_y = \frac{1}{3} y + 25 = 0 $$
$$ \frac{1}{3} y = -25 $$
$$ y = -75 $$
Since the number of units produced cannot be negative, the minimum cost occurs when 21 units are produced in Los Angeles and 75 units are produced in Denver.

# 5.

Evaluate the double integral on the given region.
$$ \int\int_R (e^{8x + 3y}) \, dA ; R: 2 \leq x \leq 4 \text{ and } 2 \leq y \leq 4 $$

Write your answer in exact form without decimals.

$$ f_x = 24 - 6y^2 = 0 $$ $$ 24 = 6y^2 $$ $$ y^2 = 4 $$ $$ y = \pm 2 $$

$$ f_y = -12xy - 24y^2 = 0 $$ $$ -12xy - 24y^2 = 0 $$ $$ -12x - 24y = 0 $$ $$ -12x = 24y $$ $$ x = -2y $$ $$ x = -2(\pm 2) $$ $$ x = \pm 4 $$

5)  Evaluate the double integral on the given region. Write your answer in exact form without decimals. $$
    \int_{2}^{4}\int_{2}^{4} e^{8x+3y} \, dx \, dy
    $$

Apply u-substitution

$u = 8x+3y$

Take derivative with respect to $x$:

$$
8+0
$$

$\frac{du}{dx} = 8+0$

$dx = \frac{1}{8}du$

Now plug in 2 and 4 to $u$ for adjusted boundaries:

$8(2)+3y=16+3y$

$8(4)+3y=32+3y$

The updated double integral becomes:

$$
\int_{16+3y}^{32+3y} e^u \cdot \frac{1}{8} du
$$

Move the constant to the front:

$$
\frac{1}{8} \int_{16+3y}^{32+3y} e^u \, du
$$

$$
\frac{1}{8}(e^{32+3y} - e^{16+3y}) \, du
$$

$$
\int_{2}^{4} \frac{1}{8}(e^{32+3y} - e^{16+3y}) \, dy
$$

Apply sum rule

$$
\frac{1}{8} \int_{2}^{4} e^{32+3y} - \int_{2}^{4} e^{16+3y} \, dy
$$

Apply u-substitution

$u = 32+3y$

Take derivative with respect to $y$:

$$
0+3
$$

$\frac{du}{dy} = 0+3$

$dy = \frac{1}{3}du$

Now plug in 2 and 4 to $u$ for adjusted boundaries:

$32+3(2)=38$

$32+3(4)=44$

The updated double integral becomes:

$$
\int_{38}^{44} e^u \cdot \frac{1}{3} \, du
$$

$$
\frac{1}{3}(e^{44}-e^{38})
$$

Apply u-substitution

$u = 16+3y$

Take derivative with respect to $y$:

$$
0+3
$$

$\frac{du}{dy} = 0+3$

$dy = \frac{1}{3}du$

Now plug in 2 and 4 to $u$ for adjusted boundaries:

$16+3(2)=22$

$16+3(4)=28$

The updated double integral becomes:

$$
\int_{22}^{28} e^u \cdot \frac{1}{3} \, du
$$

$$
\frac{1}{3}(e^{28}-e^{22})
$$

Now combine:

$$
(e^{44}-e^{38})-(e^{28}-e^{22})
$$

$$
e^{44}-e^{38}-e^{28}+e^{22}
$$

Insert the constants:

$$
\frac{1}{8} \cdot \frac13(e^{44}-e^{38}-e^{28}+e^{22})
$$

$$
\frac{1}{24}(e^{44}-e^{38}-e^{28}+e^{22})
$$

$$
\frac{e^{44}-e^{38}-e^{28}+e^{22}}{24}
$$